How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-01 tse-paper/content.tex
auto-wrap my old text for consistency

\subsection{Augmentation of Build Logs}
\label{sec:rw-bl-analysis}
Build logs are a valuable data source for developers to find out why their build failed.
Several researchers investigate how to support developers in working with verbose build logs.
Vassallo et al.~\cite{vassallo2018un-break} try to shorten the time it takes developers to understand build logs.
They parse Maven~\cite{maven2019website} build logs into a structured representation and create hint generators.
%The hint generators leverage this structured access to the information within the build log to propose fixes.
For example, one of the hint generators queries stack overflow for discussions related to why the build failed.
In a qualitative study they observed that highlighting the locality and context of an issue is helpful to programmers.
%Their tool BART is published as a Jenkins Plugin~\cite{bart2019plugin}.
The chunk retrieval techniques we compare in this article can be used to fill similar structured representations with information from build logs.
As they simplify the construction of parsers they would enable developers and researchers to cover a wider array of build tools, which is the main influence factor on the structure of a build log.

@@

The first goal in analyzing system log files is ubiquitously to separate constant and variable parts within a log message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log messages are clustered into log events, unifying messages with identical constant parts and varying parameters.
The output of a log parser is an ordered list of timed events and their corresponding parameter values~\cite{he2016evaluation}.
This structured log is then the input to various machine learning and data mining processes.
%Researchers mine patterns for operational profiling~\cite{nagappan2009efficiently}, debugging~\cite{oliner2012advances}, performance analytics or anomaly detection~\cite{nagappan2010abstracting}.
%Xu et al.~\cite{xu2009detecting} leverage the connection of log statements to the source code producing them to separate messages into constant and variable parts more accurately.

The techniques developed for system log analysis can also be applied to build logs.
One example is comparing execution traces to reference traces of intended behavior to detect anomalies.
Amar et al.~\cite{amar2019mining} employed a similar approach to detect relevant lines in build logs.
%Classic log parsers interpret the whole log file into a sequence of events.
%A similar approach could also be applied to build logs to determine the sequence of executed build steps or phases.
In this article, we focus on extracting a single specified information from the build log as a whole with chunk retrieval techniques.
Chunk retrieval techniques are used as a part of log parsing to retrieve the values of variable parts in a log message, e.g. by using regular expressions~\cite{nagappan2010abstracting,xu2009detecting}.

@@

\subsection{Program Synthesis by Example}
\label{sec:rw-prose}
% \section{Information Extraction and Retrieval Techniques}
%The techniques we investigate are based on existing methods of Programming by Example.%, information extraction and information retrieval.
%This section presents different Programming by Example resources surrounding the PROSE library.
%We explain its generic program synthesis algorithm and how PROSE synthesizes text extraction programs, the foundation of PBE\@.

@@

% In addition to I/O examples of the intended program, the user can also provide examples with only input or negative input examples.
% Negative input examples should not be processed by the synthesized program.

% The different applications of PROSE presented in the following paragraphs were all eventually implemented as DSLs for the FlashMeta synthesis algorithm.

\subsubsection{FlashExtract: Data Extraction by Example}
Le et al.~\cite{le2014flashextract:} developed FlashExtract as a DSL for the Microsoft PROSE framework.
% It enables a user to define text extraction programs for text, websites and spreadsheets by giving I/O examples.
FlashExtract's instantiation for text synthesizes extraction programs from semi-structured text based on regular expressions.
Users can extract multiple fields and structure them with hierarchy and sequence.
% FlashExtract synthesizes programs to extract each of the fields leveraging the information about hierarchical containment and sequentiality.
% It eliminates the need for the user to understand the entire structure of the processed document and decreases the effort of developing a suitable extraction program.

@@

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
% Text Similarity approaches are used to filter unstructured textual software artifacts~\cite{runeson2007detection,marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
% One common and simple technique is the Vector Space Model.
% We investigate when text similarity is a suitable technique to retrieve information chunks from build logs.
% In the following we will explain the concept of how we apply text similarity to information retrieval from CI build logs, which we refer to as \emph{CTS}\@.

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a large amount of unstructured information, a first ad-hoc approach they use is to search for related keywords.
Indeed, this was one of the most common approaches we took when searching for the reason the build failed within a log while creating the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
% As this is a technique readily available in many tools developers use to view build logs, we study when such a keyword search is suitable for retrieving information chunks from CI build logs.
% In the following we will explain how we use simple keyword search to retrieve information from CI build logs, which we refer to as \emph{KWS}\@.

@@

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
% \emph{Programming by Example} is a technique which synthesizes programs according
% to in- and output examples provided by the user.
% It enables users to create programs without a priori programming knowledge~\cite{mayer2015user}.
% In the context of text extraction through regular expressions, Programming by Example relieves the developer from having to understand the whole document structure to solve a single extraction task~\cite{le2014flashextract:}.
% In this work, we refer to our interpretation of Programming by Example as \emph{PBE}\@.
% We investigate the suitability of PBE to retrieve information chunks from build logs.
% In the following, we explain the configuration and application of PBE to chunk retrieval from CI build logs.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by Example, we refer to them in short
as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is an array of substrings of the log file text, representing the substrings that should be retrieved by the synthesized program when given the corresponding input file.
One or multiple examples, the training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program consistent with all training examples.
% A program is consistent with an example if it returns the defined output when executed on the defined input~\cite{mitchell1982generalization}.
PBE reports an error back to the user if it could not synthesize a consistent program.
% The program synthesis builds on the FlashExtract DSL, which in turn uses the FlashMeta algorithm.
% Both are described in Section~\ref{sec:rw-prose}.

% TODO Caro: The following is very obvious and could apply to any technique. 
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the synthesized regular expression program.
It then returns the substring of the build log matched by the program or an empty string if the program found no match.

@@

\subsection{Other Techniques}
%Literature mentions further build log analysis techniques.
%This section describes them with our notion of chunk retrieval techniques.

@@

\vspace{0.2cm}
\begin{itemize}[leftmargin=0.4cm]
	 \itemsep1em
	\item $|\mbox{True positives}| = \mathit{DesiredLines} \cap \mathit{RetrievedLines}$ \vspace{0.2cm}\\
        True positives are lines that appear both in the output of a tool as well as what \textit{LogChunks} contains for each example. 
        % TODO (Caro) What about when a line is replicated twice? I.e., Are line numbers part of this?
	\item $\mbox{Precision} = \dfrac{|\mathit{True\ Positives}|}{|\mathit{RetrievedLines}|}$ \vspace{0.21cm} \\
        Precision of a chunk retrieval describes which proportion of the retrieved lines were actually desired. 
	\item $\mbox{Recall} = \dfrac{|\mathit{TruePositives}|}{|\mathit{DesiredLines}|}$ \vspace{0.2cm} \\
        Recall of a chunk retrieval describes which proportion of the desired lines were retrieved.
	\item $\mbox{F$_{1}$-score} = 2 \cdot \dfrac{\mathit{Precision} \cdot \mathit{Recall}}{\mathit{Precision} + \mathit{Recall}}$ \vspace{0.2cm}\\
        In addition, we calculate the F$_{1}$-score, the harmonic mean of precision and recall. We prefer F$_{1}$ to other aggregate measures such as accuracy because for a ``needle-in-the-haystack'' scenario, we do not want to bloat our results by correctly not finding lots of irrelevant log lines.
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\ \mathit{Recall} = 1, \mathit{false\ otherwise}$  \vspace{0.2cm} \\ We define a successful retrieval as one where all desired lines were extracted, therefore when recall is one.
\end{itemize}

% Recall and precision of CTS and KWS vary with the number of lines selected for retrieval.
% We evaluate the effect of varying the number of extracted lines by multiplying the average number of lines present in the training examples with a \emph{retrieval size factor} from 0.5 to 2.5 in steps of 0.5.

@@

% Figure~\ref{fig:contextsizefactor-precision-recall-KWS} shows the effect of the retrieval size factor on precision, recall and F$_{1}$-score of chunk retrieval runs with KWS\@.
% The precision is 19\% when retrieving half of the expected number of lines. On average 9\% of the lines in the build log are retrieved then.
% When 2.5 times the expected number of lines are retrieved, the precision decreases to 10\% and a quarter of the lines in the build log are retrieved on average.
% The recall ranges from 58\% to 75\% and the F$_{1}$-score shows a constant decrease from 29\% to 17\%.

@@

% The averaged precision, recall and F$_{1}$-score f all techniques is compared in Figure~\ref{fig:recall-precision-all}.
% The recall of PBE has a high skew towards one and zero, meaning in most cases either the retrieval is successful or no relevant lines are extracted at all.
% PBE has the highest average precision with 95\%.
% Chunk retrieval with CTS has the highest average F$_{1}$-score with 51\% and the second highest recall with 46\%.
% KWS has the smallest precision of the three chunk retrieval techniques.
% With 16\% it is still higher than the precision of the RLR baseline with 7\%.
% KWS has the highest recall of all techniques with 70\%.

@@

\noindent
\textbf{Implementation}
\label{sec:prose:impl}
% TODO we need to say something about why we think that these limtiations do not actually harm our results
Our results depend on the implementation of the investigated chunk retrieval techniques and the libraries we used.
Our implementation of PBE is based on the program synthesis provided by PROSE\@.
The idiosyncrazies of this framework influence our PBE results. Other frameworks similar to PROSE might have somewhat different strengths and weaknesses.
For example, the need for examples from a single structural category stems from the fact that PROSE cannot learn regular expression programs with arbitrary boolean conjunctions and disjunctions~\cite{mayer2015user}. PROSE introduced this constraint to keep the synthesis performance reasonable. At the same time, this is clearly a current theoretical challenge of all PBE implementations, and we therefore attribute it to the technique itself, rather than the specific implementation.

@@

\noindent
\textbf{Few Samples with Many Structural Categories}
Our comparison study shows fewer measurements with many structural categories than with one category.
% TODO Caro: Please explain the following further. Why does a realistic data set only have one or few structural categories?
This stems from the fact that we use a realistic data set which in many cases has only one or few structural categories.

>>>>

\subsection{Augmentation of Build Logs}
\label{sec:rw-bl-analysis}
Build logs are a valuable data source for developers to find out why
their build failed. Several researchers investigate how to support
developers in working with verbose build logs. Vassallo et
al.~\cite{vassallo2018un-break} try to shorten the time it takes
developers to understand build logs. They parse
Maven~\cite{maven2019website} build logs into a structured
representation and create hint generators.
%The hint generators leverage this structured access to the
%information within the build log to propose fixes.
For example, one of the hint generators queries stack overflow for
discussions related to why the build failed. In a qualitative study
they observed that highlighting the locality and context of an issue
is helpful to programmers.
%Their tool BART is published as a Jenkins
%Plugin~\cite{bart2019plugin}.
The chunk retrieval techniques we compare in this article can be used
to fill similar structured representations with information from build
logs. As they simplify the construction of parsers they would enable
developers and researchers to cover a wider array of build tools,
which is the main influence factor on the structure of a build log.

@@

The first goal in analyzing system log files is ubiquitously to
separate constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}. Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters. The output of a log
parser is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}. This structured log is then
the input to various machine learning and data mining processes.
%Researchers mine patterns for operational
%profiling~\cite{nagappan2009efficiently},
%debugging~\cite{oliner2012advances}, performance analytics or anomaly
%detection~\cite{nagappan2010abstracting}. Xu et
%al.~\cite{xu2009detecting} leverage the connection of log statements
%to the source code producing them to separate messages into constant
%and variable parts more accurately.

The techniques developed for system log analysis can also be applied
to build logs. One example is comparing execution traces to reference
traces of intended behavior to detect anomalies. Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
%Classic log parsers interpret the whole log file into a sequence of
%events. A similar approach could also be applied to build logs to
%determine the sequence of executed build steps or phases.
In this article, we focus on extracting a single specified information
from the build log as a whole with chunk retrieval techniques. Chunk
retrieval techniques are used as a part of log parsing to retrieve the
values of variable parts in a log message, e.g. by using regular
expressions~\cite{nagappan2010abstracting,xu2009detecting}.

@@

\subsection{Program Synthesis by Example}
\label{sec:rw-prose}
% \section{Information Extraction and Retrieval Techniques} The
% techniques we investigate are based on existing methods of
% Programming by Example.%, information extraction and information
% retrieval. This section presents different Programming by Example
% resources surrounding the PROSE library. We explain its generic
% program synthesis algorithm and how PROSE synthesizes text
% extraction programs, the foundation of PBE\@.

@@

% In addition to I/O examples of the intended program, the user can
% also provide examples with only input or negative input examples.
% Negative input examples should not be processed by the synthesized
% program.

% The different applications of PROSE presented in the following
% paragraphs were all eventually implemented as DSLs for the FlashMeta
% synthesis algorithm.

\subsubsection{FlashExtract: Data Extraction by Example}
Le et al.~\cite{le2014flashextract:} developed FlashExtract as a DSL
for the Microsoft PROSE framework.
% It enables a user to define text extraction programs for text,
% websites and spreadsheets by giving I/O examples.
FlashExtract's instantiation for text synthesizes extraction programs
from semi-structured text based on regular expressions. Users can
extract multiple fields and structure them with hierarchy and
sequence.
% FlashExtract synthesizes programs to extract each of the fields
% leveraging the information about hierarchical containment and
% sequentiality. It eliminates the need for the user to understand the
% entire structure of the processed document and decreases the effort
% of developing a suitable extraction program.

@@

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
% Text Similarity approaches are used to filter unstructured textual
% software
% artifacts~\cite{runeson2007detection,marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
% One common and simple technique is the Vector Space Model. We
% investigate when text similarity is a suitable technique to retrieve
% information chunks from build logs. In the following we will explain
% the concept of how we apply text similarity to information retrieval
% from CI build logs, which we refer to as \emph{CTS}\@.

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad-hoc approach they
use is to search for related keywords. Indeed, this was one of the
most common approaches we took when searching for the reason the build
failed within a log while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.
% As this is a technique readily available in many tools developers
% use to view build logs, we study when such a keyword search is
% suitable for retrieving information chunks from CI build logs. In
% the following we will explain how we use simple keyword search to
% retrieve information from CI build logs, which we refer to as
% \emph{KWS}\@.

@@

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
% \emph{Programming by Example} is a technique which synthesizes
% programs according to in- and output examples provided by the user.
% It enables users to create programs without a priori programming
% knowledge~\cite{mayer2015user}. In the context of text extraction
% through regular expressions, Programming by Example relieves the
% developer from having to understand the whole document structure to
% solve a single extraction task~\cite{le2014flashextract:}. In this
% work, we refer to our interpretation of Programming by Example as
% \emph{PBE}\@. We investigate the suitability of PBE to retrieve
% information chunks from build logs. In the following, we explain the
% configuration and application of PBE to chunk retrieval from CI
% build logs.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example, we refer to them in short as \emph{examples}. The
\emph{input} is the text of the build log file. The \emph{output} is
an array of substrings of the log file text, representing the
substrings that should be retrieved by the synthesized program when
given the corresponding input file. One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted. The
PROSE program synthesis then tries to construct a program consistent
with all training examples.
% A program is consistent with an example if it returns the defined
% output when executed on the defined
% input~\cite{mitchell1982generalization}.
PBE reports an error back to the user if it could not synthesize a
consistent program.
% The program synthesis builds on the FlashExtract DSL, which in turn
% uses the FlashMeta algorithm. Both are described in
% Section~\ref{sec:rw-prose}.

% TODO Caro: The following is very obvious and could apply to any
% technique. 
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program. It then returns the substring
of the build log matched by the program or an empty string if the
program found no match.

@@

\subsection{Other Techniques}
%Literature mentions further build log analysis techniques. This
%section describes them with our notion of chunk retrieval techniques.

@@

\vspace{0.2cm}
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
	\item $|\mbox{True positives}| = \mathit{DesiredLines} \cap
	\mathit{RetrievedLines}$ \vspace{0.2cm}\\
        True positives are lines that appear both in the output of a
        tool as well as what \textit{LogChunks} contains for each
        example. 
        % TODO (Caro) What about when a line is replicated twice?
        % I.e., Are line numbers part of this?
	\item $\mbox{Precision} = \dfrac{|\mathit{True\
	Positives}|}{|\mathit{RetrievedLines}|}$ \vspace{0.21cm} \\
        Precision of a chunk retrieval describes which proportion of
        the retrieved lines were actually desired. 
	\item $\mbox{Recall} =
	\dfrac{|\mathit{TruePositives}|}{|\mathit{DesiredLines}|}$
	\vspace{0.2cm} \\
        Recall of a chunk retrieval describes which proportion of the
        desired lines were retrieved.
	\item $\mbox{F$_{1}$-score} = 2 \cdot \dfrac{\mathit{Precision}
	\cdot \mathit{Recall}}{\mathit{Precision} + \mathit{Recall}}$
	\vspace{0.2cm}\\
        In addition, we calculate the F$_{1}$-score, the harmonic mean
        of precision and recall. We prefer F$_{1}$ to other aggregate
        measures such as accuracy because for a
        ``needle-in-the-haystack'' scenario, we do not want to bloat
        our results by correctly not finding lots of irrelevant log
        lines.
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}$  \vspace{0.2cm} \\
	We define a successful retrieval as one where all desired lines were
	extracted, therefore when recall is one.
\end{itemize}

% Recall and precision of CTS and KWS vary with the number of lines
% selected for retrieval. We evaluate the effect of varying the number
% of extracted lines by multiplying the average number of lines
% present in the training examples with a \emph{retrieval size factor}
% from 0.5 to 2.5 in steps of 0.5.

@@

% Figure~\ref{fig:contextsizefactor-precision-recall-KWS} shows the
% effect of the retrieval size factor on precision, recall and
% F$_{1}$-score of chunk retrieval runs with KWS\@. The precision is
% 19\% when retrieving half of the expected number of lines. On
% average 9\% of the lines in the build log are retrieved then. When
% 2.5 times the expected number of lines are retrieved, the precision
% decreases to 10\% and a quarter of the lines in the build log are
% retrieved on average. The recall ranges from 58\% to 75\% and the
% F$_{1}$-score shows a constant decrease from 29\% to 17\%.

@@

% The averaged precision, recall and F$_{1}$-score f all techniques is
% compared in Figure~\ref{fig:recall-precision-all}. The recall of PBE
% has a high skew towards one and zero, meaning in most cases either
% the retrieval is successful or no relevant lines are extracted at
% all. PBE has the highest average precision with 95\%. Chunk
% retrieval with CTS has the highest average F$_{1}$-score with 51\%
% and the second highest recall with 46\%. KWS has the smallest
% precision of the three chunk retrieval techniques. With 16\% it is
% still higher than the precision of the RLR baseline with 7\%. KWS
% has the highest recall of all techniques with 70\%.

@@

\noindent
\textbf{Implementation}
\label{sec:prose:impl}
% TODO we need to say something about why we think that these
% limtiations do not actually harm our results
Our results depend on the implementation of the investigated chunk
retrieval techniques and the libraries we used. Our implementation of
PBE is based on the program synthesis provided by PROSE\@. The
idiosyncrazies of this framework influence our PBE results. Other
frameworks similar to PROSE might have somewhat different strengths
and weaknesses. For example, the need for examples from a single
structural category stems from the fact that PROSE cannot learn
regular expression programs with arbitrary boolean conjunctions and
disjunctions~\cite{mayer2015user}. PROSE introduced this constraint to
keep the synthesis performance reasonable. At the same time, this is
clearly a current theoretical challenge of all PBE implementations,
and we therefore attribute it to the technique itself, rather than the
specific implementation.

@@

\noindent
\textbf{Few Samples with Many Structural Categories}
Our comparison study shows fewer measurements with many structural
categories than with one category.
% TODO Caro: Please explain the following further. Why does a
% realistic data set only have one or few structural categories?
This stems from the fact that we use a realistic data set which in
many cases has only one or few structural categories.
